---
title: "SemanticDistance Word Pairs"
author: "Jamie Reilly, Hannah Mechtanberg, Emily Myers, Jonathan E. Peelle"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{SemanticDistance Word Pairs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output: 
  html_document:
    toc: true
---


```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r, message=FALSE, echo=F, warning=F}
# Load SemanticDistance
library(SemanticDistance)
```

# Word Pairs
Sample dataframe included in package. Word pairs are arrayed in columns. Columns need not be immediately adjacent within your dataframe.
```{r, eval=T, message=F, warning=F, echo=F}
knitr::kable(head(Word_Pairs, 5), format = "simple")
```

## Clean Word Pairs in Columns Transcript 
Arguments to `clean_paired_cols` are:  <br>
`dat` your raw dataframe with two columns of paired text <br>
`word1` quoted variable reflecting the column name where your first word lives <br>
`word2` quoted variable reflecting the column name where your first word lives <br>
`clean` applies cleaning functions, T/F default is TRUE <br>
`omit_stops` omits stopwords, T/F default is TRUE <br>
`lemmatize` transforms raw word to lemmatized form, T/F default is TRUE 
```{r, message=FALSE}
WordPairs_Clean <- clean_paired_cols(dat=Word_Pairs, wordcol1='word1', wordcol2='word2', clean=TRUE, omit_stops=TRUE, lemmatize=TRUE)
knitr::kable(head(WordPairs_Clean, 6), format = "simple", digits=2) 
```

## Word Pairs Semantic Distance 
Generates semantic distances (Glove and SD15) between word pairs in separate columns. Output of 'dist_paired_cols' on 2-column arrayed dataframe. Argument to `dist_paired_cols`: 
`dat` = dataframe with word pairs arrayed in columns cleaned and prepped using 'clean_2cols' fn
```{r, message=FALSE}
Columns_Dists <- dist_paired_cols(dat=WordPairs_Clean) 
knitr::kable(head(Columns_Dists, 6), format = "simple", digits=2) 
```

