---
title: "SemanticDistance Unordered List"
author: "Jamie Reilly, Hannah Mechtenberg, Emily Myers, Jonathan E. Peelle"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{SemanticDistance Unordered List}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
output: 
  html_document:
    toc: true
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r, message=FALSE, echo=F, warning=F}
# Load SemanticDistance
library(SemanticDistance)
```

# Unordered Lists
An unordered word list is just a bag-of-words. You might be interested in applying kmeans, hierarchical clustering, or graph metrics to elucidate a structure. In all of these instances, you will pass R a vector of words in no particular order. Here's a sample dataframe with exemplars by semantic category. Let's see if this function can find order in chaos!

<span style="color: darkred;">Sample unordered word list included in the package</span> 
```{r, eval=T, message=F, warning=F}
#print first ten rows of header
knitr::kable(head(Unordered_List, 1), format = "pipe")
```

## Clean Unordered List 
Prep a vector of words for hierarchical clustering or a network visualization. `clean_unordered` will retain only one instance of each string (distinct, no duplicates) and no missing values. <br>
Arguments to `clean_unordered` are: <br>
`dat` raw dataframe with at least one column of text <br>
`wordcol` quoted variable reflecting where your text lives <br>
`clean` applies cleaning functions, T/F default is TRUE <br>
`omit_stops` omits stopwords, T/F default is TRUE <br>
`lemmatize` transforms raw word to lemmatized form, T/F default is TRUE <br>
```{r, message=FALSE}
Clusters_Clean <- clean_unordered(dat=Unordered_List, wordcol="mytext", clean=TRUE, omit_stops=TRUE, lemmatize=TRUE)
head(Clusters_Clean)
```

## Convert cleaned list to square matrix
On our road to deriving a cluster analysis or network, we need to compute distances between all pairwise elements. This involves creating a square matrix where each entry [i,j] is the cosine distance between word i and word j. Matrix contains original words as both row and column names for reference. User specifies whether to return a matrix based on embeddings (GLOVE) or experiential norms (SD15). Returns square matrix where each entry [i,j] is the cosine distance between word i and word j. Matrix contains original words as both row and column names for reference. Arguments to `dist_matrix` are: <br>
`dat` dataframe cleaned and prepped using 'clean_unordered4matrix' fn <br>
`dist_type` quoted argument default is 'embedding', other option is "SD15"
```{r, message=FALSE}
MyDistMatrix <- dist_matrix(dat=Clusters_Clean, dist_type='embedding')
MyDistMatrix[1:7, 1:7] #Print columns 1:7, rows 1:7 square matrix
```

## Step 2: Matrix to Dendrogram
```{r, message=FALSE, warning=FALSE}
str(MyDistMatrix)
MyDendro <- viz_clusters(MyDistMatrix, type="dendrogram")
# plots triangular dendrogram with K clusters
```

## Step 3: Matrix to iGraph Network
```{r, message=FALSE, warning=FALSE}
MyGraph <- viz_clusters(MyDistMatrix, type="network")
```

