


```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

```{r, message=FALSE, echo=F, warning=F}
# Load SemanticDistance
library(SemanticDistance)
```

# Dialogues
This could be a conversation transcript or any language sample where you care about talker/interlocutor information (e.g., computing semantic distance across turns in a conversation). Your dataframe should nominally contain a text column and a speaker/talker column. 

<span style="color: darkred;">sample dialogue transcript included in the package</span> 
```{r, eval=T, message=F, warning=F}
MaryLittleLamb <- read_1file(MaronGross_2013)
#print first ten rows of header
knitr::kable(head(MaronGross_2013, 10), format = "pipe")
```

## <span style="color: darkred;">Step 1: Clean Dialogue Transcript (clean_dialogue) </span> <br>
Decide on your cleaning parameters (e.g., stopwords? lemmatization?). Specify these in the argument(s) to your function calls.

**Arguments to 'clean_dialogue' are:** <br>
**dat** = your raw dataframe with at least one column of text AND a talker column <br>
**wordcol** = column name (quoted) containing the text you want cleaned <br>
**whotalks** = column name (quoted) containing the talker ID (will convert to factor) <br>
**clean** = applies cleaning function, T/F default is TRUE <br>
**omit_stops** = omits stopwords, T/F default is TRUE <br>
**lemmatize** = transforms raw word to lemmatized form, T/F default is TRUE
```{r, message=FALSE}
Dialogue_Cleaned <- clean_dialogue(dat=Dialogue_Structured, wordcol="mytext", whotalks = "speaker", clean=TRUE, omit_stops=TRUE, lemmatize=TRUE, split_strings=TRUE)
head(Dialogue_Cleaned, n=12)
```

## Step 2: Compute Semantic Distances 
## <span style="color: darkred;"> Dialogue Distance Turn-to-Turn (dist_dialogue)</span>
Averages the semantic vectors for all content words in a turn then computes the cosine distance to the average of the semantic vectors of the content words in the subsequent turn.  Note: this function only works on dialogue samples marked by a talker variable (e.g., conversation transcripts). It averages across the semantic vectors of all words within a turn and then computes cosine distance to all the words in the next turn. You just need to feed it a transcript formatted with clean_dialogue. 'dist_dialogue' will return a summary dataframe that distance values aggregated by talker and turn (id_turn). <br>
<br>

**Arguments to 'dist_dialogue' are:**  <br>
**dat** = dataframe w/ a dialogue sample cleaned and prepped using 'clean_dialogue'
```{r, message=FALSE}
DialogueDists <- dist_dialogue(dat=Dialogue_Cleaned)
head(DialogueDists)
```

